R Notebook

### 2018-05-13 analysis/final_analysis.Rmd
Do a bunch of bootstrapping for Expt. 1 (messy)

```{r}
critTrials <- exp1_d %>% 
  filter(attemptNum == 0) %>%
  filter(critical == 1) %>%
  filter(trialType == 'exp')

@@

```{r}
critTrials %>%
  group_by(condition) %>%
  mutate(error = 1 - correct) %>%
  tidyboot_mean(column = error) %>%
  ggplot(aes(x = condition, y =empirical_stat, fill = condition)) +
    geom_bar(stat = 'identity', alpha = .5) +
    geom_errorbar(aes(ymax = ci_upper, ymin = ci_lower), width = 0) +
    ylab('% errors') +
    scale_fill_colorblind()+
    theme_few() +
    theme(aspect.ratio = 1, legend.position="none")

@@

### Control for item-level error rates...

@@

```{r}
diffs <- critTrials %>% 
  mutate(error = 1 - correct) %>%
  group_by(objectSet, condition, targetObject) %>%
  summarize(pctErrors = mean(error)) %>%
  spread(condition, pctErrors) %>%
  mutate(errorReduction = scripted - unscripted) %>%
  gather(condition, pctErrors, scripted, unscripted)

critTrials %>% 
  left_join(diffs) %>%
  mutate(error = 1 - correct) %>%
  group_by(objectSet, condition, targetObject, errorReduction) %>%
  tidyboot_mean(column = error) %>%
  ggplot(aes(x = reorder(targetObject, -errorReduction), y = empirical_stat, color = condition, group = condition)) +
    geom_point() +
    geom_line(stat = 'identity', position = 'dodge') +
    geom_errorbar(aes(ymax = ci_upper, ymin = ci_lower), width = 0) +
    theme_few() +
    theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
    ggtitle('error heterogeneity across item') +
    xlab('critical item') +
    ylab('% error in condition')

@@

```{r}
summary(glmer(hovered ~ trialType * condition + (1 + trialType|gameid) + (1 |objectSet), data = d.onDistractor, family = 'binomial'))
```

@@

For 6 of the 8 items, unscripted speakers had significantly more words in their referring expressions... (n.s. for the other 2, but idiosyncratic)

```{r}
scripted_refs <- c('glasses', 'bottom block', 'tape', 'large measuring cup', 'brush', 'eraser', 'small candle', 'mouse')
standardized_refs <- read_csv('../data/final/experiment1/refExpressions_standardized.csv')
length(unique(c(standardized_refs$text, scripted_refs)))
unscripted_refs <- read_csv('../data/final/experiment1/refExpressions.csv')
scripted.df <- data.frame(objectSet = c(1,2,3,4,5,6,7,8), m = scripted_refs, numRawWords = c(1,2,1,3,1,1,2,1))
refs %>% 
  mutate(numRawWords = str_count(text, "\\S+")) %>%
  group_by(objectSet) %>%
  tidyboot_mean(numRawWords) %>%
  left_join(scripted)
```

@@

```{r}
dodge <- position_dodge(width=0.9)

d %>%
  group_by(context, hidden) %>%
  tidyboot_mean(column = numRawWords) %>%
  ggplot(aes(x = context, y = empirical_stat, fill = hidden)) +
    geom_bar(stat = 'identity', position = dodge) +
    geom_errorbar(aes(ymin = ci_lower, ymax = ci_upper), position = dodge, width = 0) +
    theme_few() +
    ylab('mean # words') +
    theme(aspect.ratio = 1, text = element_text(size=20)) +
    scale_fill_colorblind()

ggsave('~/Downloads/aggregated_means.pdf', width = 4, height = 2)
```

@@

```{r}
hidden.d %>% 
  filter(context == 'far') %>% 
  mutate(group = case_when(diff < -0 ~ 'anti',
                           diff > 0 ~ 'pro',
                           TRUE ~ 'all_the_things')) %>%
  group_by(group) %>% tally()

@@

@@

ggsave('~/Downloads/scatter.pdf')
```

@@

ggplot(d.booted, aes(x = context, y = empirical_stat, fill = hidden)) + 
  geom_bar(stat = 'identity', position = dodge) +
  geom_errorbar(aes(ymax = ci_upper, ymin = ci_lower), position = dodge, width = 0) +
  facet_wrap(~ feature) + 
  theme_few() +
  theme(aspect.ratio = 1, text = element_text(size=20))  +
  ylab('% utterances with feature') +
  scale_fill_colorblind()

ggsave('~/Downloads/by-feature-analysis.pdf')
```

>>>>

```{r}
critTrials <- exp1_d %>% 
  filter(attemptNum == 0) %>%
  filter(critical == 1) %>%
  filter(trialType == 'exp') %>%
  mutate(error = 1 - correct)

@@

```{r}
critTrials %>%
  group_by(condition) %>%
  tidyboot_mean(column = error) %>%
  ggplot(aes(x = condition, y =empirical_stat, fill = condition)) +
    geom_bar(stat = 'identity', alpha = .5) +
    geom_errorbar(aes(ymax = ci_upper, ymin = ci_lower), width = 0) +
    ylab('% errors') +
    scale_fill_colorblind()+
    theme_few() +
    theme(aspect.ratio = 1, legend.position="none")

@@

### Control for item-level and subject-level error rates...

@@

```{r}
error_diffs <- critTrials %>% 
  group_by(objectSet, condition, targetObject) %>%
  summarize(pctErrors = mean(error)) %>%
  spread(condition, pctErrors) %>%
  mutate(errorReduction = scripted - unscripted) %>%
  gather(condition, pctErrors, scripted, unscripted) %>%
  right_join(critTrials) %>%
  group_by(objectSet, condition, targetObject, errorReduction) %>%
  tidyboot_mean(column = error)

ggplot(error_diffs, aes(x = reorder(targetObject, -errorReduction), y = empirical_stat, color = condition, group = condition)) +
    geom_point() +
    geom_line(stat = 'identity', position = 'dodge') +
    geom_errorbar(aes(ymax = ci_upper, ymin = ci_lower), width = 0) +
    theme_few() +
    theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
    ggtitle('error heterogeneity across item') +
    xlab('critical item') +
    ylab('% error in condition')

@@

```{r}
summary(glmer(hovered ~ trialType * condition + (1 + trialType |gameid) + (1 |objectSet), data = d.onDistractor, family = 'binomial'))
```
Note that this is still significant for our preregistered dv (raw amount of data) -- it's just less interpretable given the distribution...

```{r}
summary(lmer(totalAmtTimeOnDistractor ~ trialType * condition + (1 + trialType|gameid) + (1 |objectSet), 
             data = d.onDistractor %>% mutate(totalAmtTimeOnDistractor = ifelse(totalAmtTimeOnDistractor == 0, 0, log(totalAmtTimeOnDistractor)))))
```

@@

Look at all of them to make sure there aren't dumb duplicates...

```{r}
scripted_words <- c('glasses', 'bottom block', 'tape', 'large measuring cup', 'brush', 'eraser', 'small candle', 'mouse')
scripted_refs <- data.frame(objectSet = c(1,2,3,4,5,6,7,8), text = scripted_words, stringsAsFactors=FALSE)

standardized_refs <- read_csv('../data/final/experiment1/refExpressions_standardized.csv')
cat('unique utterances: ', length(unique(c(scripted_refs, standardized_refs$text))))
unscripted_utt_counts <- standardized_refs %>% 
  group_by(objectSet,text) %>%
  tally() %>%
  mutate(condition = 'unscripted')
```

Unscripted speakers had significantly more words in their referring expressions across items... (note: if you just bootstrap cis for each item, it's significant for 6 of 8...)

```{r}
unscripted_refs <- read_csv('../data/final/experiment1/refExpressions.csv')
scripted.df <- data.frame(objectSet = c(1,2,3,4,5,6,7,8), scriptedMessage = scripted_refs, scriptedNumWords = c(1,2,1,3,1,1,2,1))
diff.df <- unscripted_refs %>% 
  mutate(unscriptedNumWords = str_count(text, "\\S+")) %>%
  left_join(scripted.df, by = c('objectSet')) %>%
  mutate(diff = unscriptedNumWords - scriptedNumWords)

library(lmerTest)
summary(lmer(diff ~ 1 + (1 | objectSet) + (1 | gameid), data= diff.df))
```

## Analyze informativity ratings

First identify 4 turkers who were responding randomly...

```{r}
raw_inf_ratings <- read_csv('../informativity_ratings.csv')
# ggplot(raw_inf_ratings %>% mutate(workerid = substr(workerid, 0, 6)), aes(x = response, fill = referent)) +
#   geom_density(alpha = 0.5) +
#   facet_wrap(~ workerid, scales = 'free_y')
# 
# ggplot(raw_inf_ratings %>% mutate(workerid = substr(workerid, 0, 6)), aes(x = slideNumber, y = response, color = referent)) + 
#   geom_point() +
#   facet_wrap(~ workerid)

raw_inf_ratings %>% mutate(workerid = substr(workerid, 0, 6)) %>%
  group_by(workerid) %>%
  summarize(medianRT = median(rt))

bad_raters <- c('956e201e2604ead2f2fa81d0ef666626', '6559a18fae791987122350d706538e90', '2301c2d66270fecb3c16cd1593e4aa0f', 'dd0463e01b62413231a9218253908a91')#, 'deec4028cd9f327085a947bac0e19aab', '766e6aa74bc2c113e29531dab30d63e9')
```

```{r}
library(irr)
inf_ratings <- raw_inf_ratings %>%
  filter(!(workerid %in% bad_raters))

inf_diffs <- inf_ratings %>%
  select(-slideNumber, -rt) %>%
  group_by(workerid, label) %>%
  spread(referent, response) %>%
  mutate(diff = target - distractor) %>%
  ungroup()

wideRatings <- inf_diffs %>% select(label, diff, workerid) %>% spread(workerid, diff)
icc(wideRatings[,-c(1)],
    model = "twoway", type = "agreement")

inf_diffs_collapsed <- inf_diffs %>%
  group_by(label) %>%
  summarize(distractor = mean(distractor),
            target = mean(target),
            paired_inf = mean(diff))
```

lmer mixed-model here is complicated because of the complex & partially crossed nature of the variance in the data (i.e. between judgements of raters, between object sets, between utterances chosen by different speakers)  TODO: double-bootstrapping thing where you first resample 'mean' ratings for each utterance, then you resample the distribution of utterances in a condition (i.e. within object set), then you take a weighted average to get your 'informativity'

```{r}
## Make a df with each utterance duplicated proportionally to the times it was said
utt_data_tmp <- scripted_refs %>% mutate(n = 68) %>% mutate(condition = 'scripted') %>%
  group_by(text) %>% 
  slice(rep(1:n(), each = n)) %>%
  mutate(gameid = as.character(row_number()))

utt_data <- left_join(
  inf_diffs %>% group_by(label) %>% summarize(m = mean(diff)),
  utt_data_tmp %>% rename(label = text)
) %>% filter(condition == 'scripted') %>% group_by(label, objectSet, condition) %>% summarize(m = mean(m)) %>%
  right_join(critTrials %>% filter(condition == 'scripted'), by = c('objectSet', 'condition')) %>%
  group_by(gameid, condition, objectSet, label) %>%
  group_by(label, objectSet, condition, gameid) %>%
  tally() %>%
  select(-n) %>%
  rename(text = label) %>%
  bind_rows(standardized_refs %>% mutate(condition = 'unscripted', objectSet = as.numeric(objectSet)))
```

```{r}
## In the inner loop, we want to sample a 'mean' rating for each utterance
## using same sample of *raters* for all utterances... 
sampleRatings = function(df, indices) {
  grouped_df <- df %>%
    group_by(label, objectSet) %>%
    do(data.frame(diff = mean(.$diff[indices]),
                  target = mean(.$target[indices]),
                  distractor = mean(.$distractor[indices])))
  return(grouped_df)
}

## In outer loop, we want to sample 'pairs' as a unit (with their 8 utterances)
samplePairs = function(df) {
  uniq_ids <- unique(df$gameid)
  new_ids <- sample(uniq_ids, length(uniq_ids), replace = TRUE)
  resampled_ids = data.frame(gameid = new_ids,
                             sampleid = 1:length(new_ids), 
                             stringsAsFactors = FALSE)
  return(left_join(resampled_ids, df, by = c('gameid')))
}

## Finally, want to sample object sets per pair
sampleObjects = function(df, indices) {
  new_objects <- data.frame(objectSet = df$objectSet[indices])
  return(left_join(new_objects, df, by = c('objectSet')))
}
# sampleObjects(data.frame(objectSet = c(1,2,3,4,5,6,7,8), utt = c('a', 'b', 'c', 'a', 'b', 'c', 'd', 'e')),
#               obj_sets)
```

```{r}
## do a multi-stage bootstrap...
## 1. resample a set of raters to get a mean for each utterance...
## 2. sample object set indices (e.g. 5,2,3,3,5,1,8,5)
## 3. resample pairs within each condition and then resample utterances within pairs using object set indices
boot_samples = replicate(1000, {
  raters = sample(1:16,16,replace = TRUE)
  obj_sets = sample(1:8, 8, replace = TRUE)
  resampled_ratings = sampleRatings(inf_diffs, raters)

  resampled_pairs = utt_data %>%
    group_by(condition) %>%
    do(samplePairs(.)) %>%
    group_by(condition, sampleid) %>%
    do(sampleObjects(., obj_sets)) %>%
    rename(label = text)

  combined <- left_join(resampled_pairs, resampled_ratings, by = 'label') %>%
    gather(referent, value, target,distractor,diff) %>%
    group_by(condition, referent) %>%
    summarize(m = mean(value))
  diff_of_diffs <- combined %>% filter(referent == 'diff') %>% spread(condition, m) %>% mutate(diff_of_diffs = unscripted - scripted)
  return(c((combined %>% filter(referent != 'diff'))$m, diff_of_diffs$diff_of_diffs))
})
```

```{r}
# Look at diff of diffs (if excludes 0, significant)
data.frame(test_stat = boot_samples[5,]) %>%
  summarize(m = mean(test_stat),
            ci_upper = sort(test_stat)[length(test_stat) * 0.025],
            ci_lower = sort(test_stat)[length(test_stat) * 0.975]) 
```

```{r}
boot_stats = data.frame(
  scripted_distractor = boot_samples[1,],
  scripted_target = boot_samples[2,],
  unscripted_distractor = boot_samples[3,],
  unscripted_target = boot_samples[4,]
) %>%
  gather(condition, value) %>%
  group_by(condition) %>%
  summarize(m = mean(value),
            ci_upper = sort(value)[length(value) * 0.025],
            ci_lower = sort(value)[length(value) * 0.975]) %>%
  separate(condition, into = c('condition', 'referent'), by = '_')

dodge <- position_dodge(width=0.9)
ggplot(boot_stats, aes(x = condition, y = m, fill = condition, alpha = referent)) +
  geom_bar(stat= 'identity', position = 'dodge') +
  scale_alpha_manual(values=c(.25, .75)) +
  geom_errorbar(aes(ymax = ci_upper, ymin = ci_lower), width = 0, position = dodge) + 
  theme_few() +
  scale_fill_colorblind() +
  guides(fill = 'none') +
  ylab('utterance informativity ratings') +
  theme(aspect.ratio = 1)

ggsave('../writing/journal_manuscript/figures/informativity.pdf')
```

### How well does informativity correlate with error?

now bootstrap again but within items

```{r}
boot_objectsets = replicate(1000, {
  raters = sample(1:16,16,replace = TRUE)
  resampled_ratings = sampleRatings(inf_diffs, raters)

  resampled_pairs = utt_data %>%
    group_by(condition) %>%
    do(samplePairs(.)) %>%
    rename(label = text)

  combined <- left_join(resampled_pairs, resampled_ratings, by = c('label', 'objectSet')) %>%
    gather(referent, value, target,distractor,diff) %>%
    group_by(condition, objectSet, referent) %>%
    summarize(m = mean(value))

  return((combined %>% filter(referent == 'diff'))$m)
})
```

Kind of messy way to aggregate these, but so it goes... (replicate doesn't return a dataframe)

(note that the 'outlier' here is the scripted 'candle' trial: raters thought it was *much* more likely to refer )

```{r}
boot_objectsets_stats = data.frame(
  condition = c(rep('scripted', 8), rep('unscripted', 8)),
  m_rating = rep(0,16),
  ci_upper_rating = rep(0,16),
  ci_lower_rating = rep(0,16),
  objectSet = rep(1:8,2)
)

for (i in 1:16) {
  boot_row = boot_objectsets[i,]
  boot_objectsets_stats[i,]$m_rating = mean(boot_row)
  boot_objectsets_stats[i,]$ci_upper_rating = sort(boot_row)[length(boot_row) * 0.025]
  boot_objectsets_stats[i,]$ci_lower_rating = sort(boot_row)[length(boot_row) * 0.975]
}

left_join(boot_objectsets_stats, error_diffs, by = c('condition', 'objectSet')) %>%
  ggplot(aes(x = m_rating, y = mean, color = condition, group = 1)) +
    geom_point(size = 2, shape = 16) +
    geom_errorbar(aes(ymax = ci_upper, ymin = ci_lower), alpha = .3) +
    geom_errorbarh(aes(xmax = ci_upper_rating, xmin = ci_lower_rating), alpha =.3) +
    geom_smooth(linetype = 'dotted', method = 'lm', colour="black", se = F) +
    theme_few() +
    theme(aspect.ratio = 1, legend.position = 'top') +
    scale_color_colorblind() +
    xlab('informativity diff (target - distractor)') +
    ylab('% critical error')

ggsave('../writing/journal_manuscript/figures/informativity_predicts_errors.pdf',
       height = 3, width = 3, useDingbats=FALSE)
```

```{r}
boot_correlation = replicate(1000, {
  raters = sample(1:16,16,replace = TRUE)
  resampled_ratings = sampleRatings(inf_diffs, raters)

  resampled_pairs = utt_data %>%
    group_by(condition) %>%
    do(samplePairs(.)) %>%
    rename(label = text)

  combined <- left_join(resampled_pairs, resampled_ratings, by = c('label', 'objectSet')) %>%
    select(-target,-distractor,-label) %>%
    inner_join(critTrials, by = c('gameid', 'condition', 'objectSet')) %>%
    group_by(objectSet, condition) %>%
    summarize(inf = mean(diff), error = mean(error))

  return(cor(combined$inf, combined$error))
})

cat('r = ', mean(boot_correlation), 
    ', 95% CI = [', sort(boot_correlation)[length(boot_correlation) * 0.025],
    ', ', sort(boot_correlation)[length(boot_correlation) * 0.975],']')
```

@@

```{r}
dodge <- position_dodge(width=0.9)
cbPalette = c(rgb(0.8, .8, .8), rgb(.8, .3, .3),  rgb(0, .6, .4), rgb(.4, .4, .4))

d %>%
  group_by(context, hidden) %>%
  tidyboot_mean(column = numRawWords) %>%
  ggplot(aes(x = context, y = empirical_stat, fill = hidden)) +
    geom_bar(stat = 'identity', position = dodge) +
    geom_errorbar(aes(ymin = ci_lower, ymax = ci_upper), position = dodge, width = 0) +
    theme_few() +
    ylab('mean # words') +
    theme(aspect.ratio = 1, text = element_text(size=20)) +
    scale_fill_manual(values=cbPalette) +
    theme(aspect.ratio = 1)

ggsave('../writing/journal_manuscript/figures/num_words.pdf', width = 4, height = 3)
```

@@

hidden.d %>% 
  filter(context == 'far') %>% 
  mutate(group = case_when(diff < -0 ~ 'anti',
                           diff > 0 ~ 'pro',
                           TRUE ~ 'all_the_things')) %>%
  group_by(group) %>% tally()

@@

```{r}

@@

ggsave('../writing/journal_manuscript/figures//scatter.pdf')
```

@@

ggplot(d.booted, aes(x = context, y = empirical_stat, fill = hidden)) + 
  geom_bar(stat = 'identity', position = dodge) +
  geom_errorbar(aes(ymax = ci_upper, ymin = ci_lower), position = dodge, width = 0) +
  facet_wrap(~ feature) + 
  theme_few() +
  theme(aspect.ratio = 1, text = element_text(size=20))  +
  ylab('% utterances with feature') +
  scale_fill_manual(values=cbPalette)

ggsave('../writing/journal_manuscript/figures/by-feature-analysis.pdf', 
       height = 3, width = 9)
```
